### 2023-07-11 documentation/article03/code/integral2.cpp.tex

#include <cmath>
#include <cstddef>
#include <iomanip>
#include <iostream>
#include <mpi.h>
#include <ostream>

inline long double f(long double x)
{
	return std::sqrt(std::pow(100, 2) - std::pow(x, 2));
}

inline long double integrate(long double (*f)(long double),
double lowerLimit,
double upperLimit,
int partitions)
{
	double h = (upperLimit - lowerLimit) / partitions;
	long double sum = 0;

	int slotId, slotsAvailable;
	MPI_Comm_rank(MPI_COMM_WORLD, &slotId);
	MPI_Comm_size(MPI_COMM_WORLD, &slotsAvailable);

	int localIterations = (partitions - 2) / slotsAvailable;
	int remainingIterations = (partitions - 2) % slotsAvailable;

	int start = slotId * localIterations + std::min(slotId, remainingIterations) + 1;
	int end = start + localIterations + (slotId < remainingIterations ? 1 : 0);

	long double localSum = 0;

	for (int k = start; k < end; k++) {
		localSum += f(lowerLimit + k * h);
	}

	MPI_Reduce(&localSum, &sum, 1, MPI_LONG_DOUBLE, MPI_SUM, 0, MPI_COMM_WORLD);

	if (slotId == 0) {
		sum *= 2;
		sum += (f(lowerLimit) + f(upperLimit));
		sum *= (h / 2);
	}

	return sum;
}

inline void experiment(double lowerLimit, double upperLimit, double interval)
{
	long double result = 0;
	int partitions = upperLimit / interval;

	double start_time = MPI_Wtime();

	result = integrate(f, lowerLimit, upperLimit, partitions);

	double end_time = MPI_Wtime();
	double elapsed_time = end_time - start_time;

	int rank;
	MPI_Comm_rank(MPI_COMM_WORLD, &rank);

	if (rank == 0) {
		std::cout << std::fixed << std::setprecision(6) << interval << " \t " << partitions
		<< " \t " << elapsed_time << " \t " << result << std::endl;
	}
}

int main(int argc, char *argv[])
{
	MPI_Init(NULL, NULL);

	double lowerLimit = 0, upperLimit = 100;

	int slotId;
	MPI_Comm_rank(MPI_COMM_WORLD, &slotId);

	if (slotId == 0) {
		std::cout << "Interval \t Partitions \t Time taken \t Result" << std::endl;
	}

	experiment(lowerLimit, upperLimit, 0.000001);
	experiment(lowerLimit, upperLimit, 0.00001);
	experiment(lowerLimit, upperLimit, 0.0001);

	MPI_Finalize();
	return 0;
}